I'm trying to write an abstract for my thesis. Can you help me?
This is what I got:

Example abstracts from other papers:

In this paper we provide a quantum Monte Carlo algorithm to
solve high-dimensional Black-Scholes PDEs with correlation
for high-dimensional option pricing. The payoff function
of the option is of general form and is only required to
be continuous and piece-wise affine (CPWA), which covers
most of the relevant payoff functions used in finance.
We provide a rigorous error analysis and complexity
analysis of our algorithm. In particular, we prove
that the computational complexity of our algorithm
is bounded polynomially in the space dimension d of
the PDE and the reciprocal of the prescribed accuracy
and so demonstrate that our quantum Monte Carlo algorithm
does not suffer from the curse of dimensionality.

Grid-free Monte Carlo methods based on the walk on spheres (WoS) algorithm
solve fundamental partial differential equations (PDEs) like the Poisson
equation without discretizing the problem domain, nor approximating functions
in a finite basis. Such methods hence avoid aliasing in the solution, and evade
the many challenges of mesh generation. Yet for problems with complex geometry,
practical grid-free methods have been largely limited to basic Dirichlet boundary
conditions. This paper introduces the walk on stars (WoSt) method, which solves
linear elliptic PDEs with arbitrary mixed Neumann and Dirichlet boundary conditions.
The key insight is that one can efficiently simulate reflecting Brownian
motion by replacing the balls used by WoS with star-shaped domains; we identify
such domains by locating the closest visible point on the geometric silhouette.
Overall, WoSt retains many attractive features of other grid-free Monte Carlo methods,
such as progressive evaluation, trivial parallel implementation, and logarithmic
scaling relative to geometric complexity.

Attempts at an abstract for my thesis:

This thesis explores unbiased Monte Carlo methods for linear ODEs and parabolic PDEs.
After introducing MC techniques we develop an


This thesis introduces recursive Monte Carlo, mainly with simple examples on
ODEs. By combining existing Monte Carlo techniques we develop an
unbiased Monte Carlo algorithm for initial value problems that achieves higher order.

This thesis provides an unbiased Monte Carlo
algorithm for linear initial value
problems that achieve higher order
comparable to classical methods. The key
to this algorithm is combining existing Monte Carlo
techniques and applying them appropriately to initial
value problems.


This thesis explores unbiased Monte Carlo methods in time
specifically for linear ODEs and parabolic PDEs.
Major attention is spent on


I adjusted your version:

This thesis explores unbiased Monte Carlo methods for solving linear
ordinary differential equations (ODEs).
The proposed algorithm
capitalizes on the appropriate combination of Monte Carlo techniques
to achieve accurate and efficient solutions.
